Functional Enrichment Tests: Intro and methods comparison

### 2023-10-05 DRAFT_Funct_Enrich.Rmd
Push to do list for SB DE analysis and GO annot QC

## Using R

# What to Do with the Output

-   Visualizing results
-   Interpreting results - gather consensus advice
-   R...Mann-Whitney U Test and Fisher's Exact Test (Sam)

# Formatting Examples

## Example of table syntax

| Term A | Term B | Term C | Term D | Term E | Term F | Term G |
|:-------|:-------|:-------|:-------|:-------|:-------|:-------|
| xxx    | xxx    | xxx    | xxx    | xxx    | xxx    | xxx    |
| yyy    | yyy    | yyy    | yyy    | yyy    | yyy    | yyy    |
| zzz    | zzz    | zzz    | zzz    | zzz    | zzz    | zzz    |

## Example of how to insert an image

```{r, echo = FALSE, out.height = "40%", fig.align = "center", }

knitr::include_graphics("DRAFT_Funct_Enrich/Rivera_etal_fig.png")

## Example of how to write and print an equation

For illustrative purposes, here is a simple and famous equation

$$
E = mc^2
$$

As an example of a more complex equation, here is an equation for the t-statistic

$$
t = \frac{\overline{x_{1}}-\overline{x_{2}}}
{\sqrt{(s^2(
\frac{1}{n_{1}} + \frac{1}{n_{2}}
))}}
$$

## Example of how to cite a reference from your .bib in-text

Functional enrichment tests applied to mRNA analyses often impose significant biases [@Bleazard2015-vn].

````         
```{r setup, include = FALSE} `r ''`

# Set root directory... in this draft, ensure that you have the same path
knitr::opts_knit$set(root.dir = '~/Documents/GitHub/marineomics.github.io/')

```
````

# Let's get started

Below is an example of what should be the first, visible code chunk

```{r, results = FALSE, message = FALSE, warning = FALSE}

## Unhash and run the code below if you believe you may need to install the packages loaded below
# Edit code below to list packages that are used in this draft... at present these are just examples

if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install(c("DESeq2","edgeR","arrayQualityMetrics"))

# Load packages
library(tidyverse)

>>>>

## Comparing Fisher's Exact vs. Mann Whitney U tests

```{r}

## To-do list for Sam once he has data

# 1. Conduct differential expression analysis w/ oyster data modeling continuous vs. categorical effects of pH (small effect; continuous) and sex (large effect; categorical) on gene expression

# 2. Using example GO annotations, perform FE and MWU tests

# 3. Compare results and discuss differences... why was one effect type influenced more than the other? How does this inform what kind of test you should use?

# 4. Querying how annotation quality is influencing results
